# Portfolio assignment 1.2   Looking ahead

##  Portfolio assignment 1.2 Part 1

<div class="azure">
<h4> Looking ahead</h4>
<h3>  Answer the following questions:  </h3>

<li>  Where do I want to be in ~2 years time? </li>
<li>  How am I doing now with respect to this goal?  </li>
<li>  What would be the next skill to learn? </li>

</div>


<h3> Plan  to get a paid job in the Netherlands </h3>

To do this I need to take the following steps: 
<p> master the Dutch language, get a basic education. </p>
<p>In my final year of education I plan to do an internship at inspection company NAK ( Randweg 14,Emmeloord). Presumably I will be doing work on detecting infection in potato seeds, using qPCR and data analyse of results. </p>

<p>The current course TLSC-DSFB26V-20 is aimed at obtaining basic bio-medical and bioinformatics knowledge, which I need for your work position. <p>
</div>

<h3> What would be the next skill to learn </h3>

<li> -Genomic Data Visualization and Interpretation (Ensemble, UCSC, IGV) </li>
<li> -Visualize mutation profile patterns (GenVisR) </li>
<li> -Visualize the genomic characterization of disease and response to therapy results (GenVisR) </li>
<li> -Matplotlib & Seaborn PyData  </li>


<div class="azure pd22"> Make a planning on how to start learning this new skill.

Note the following points:

    It has to be life sciences or useful for life sciences (no knitting, riding horses etc)
    It has to be data science / bioinformatics
    You have to show your plan in your portfolio, together with the result of the first few steps you took in trying to reach this goal.
    You probably won’t learn this skill in a few days. Just get started.

This assignment will show us whether you succeeded in reaching an overview of the field and possibilities in data science for biology</div>


## Portfolio assignment 1.2 Looking ahead. Package GenVisR


<h3> Visualizing genetic Mutations  with GenVisr  </h3>

 <p> * this is the first approach. The extended analyse follows.


<div class="pd22">

```{r mutation_viz, echo=FALSE}


library(GenVisR)
library(data.table)

mutation_data <-  fread("data/mutation_data.tsv")

mutation_data <- mutation_data[,c('patient', 'gene name','trv type','amino acid change')]

setnames(mutation_data, c("sample", "gene", "mutation", "aminoacid change"))

myHierarchy <- data.table("mutation"=c("nonsense",
                                       "frame_shift_del", "frame_shift_ins", "in_frame_del",
                                       "splice_site_del", "splice_site", "missense",
                                       "splice_region", "rna"), color=c("#FF0000", "#00A08A",
                                                                        "#F2AD00", "#F98400", "#5BBCD6", "#046C9A", "#D69C4E",
                                                                        "#000000", "#446455"))

plotData <- Waterfall(mutation_data, mutationHierarchy = myHierarchy)

```

</div>



```{r waterfall_plot}


drawPlot(plotData)


```

<h3> visualization cancerogenic gene  PIK3CA  </h3>

</br>
</br>




```{r mutation_gen_PIK3CA, echo=FALSE, warning=FALSE, message=FALSE}

library("g3viz")

mutation_gene_hugo <- "PIK3CA"

mutation.csv <- system.file("extdata", "ccle.csv", package = "g3viz")


mutation.dat <- readMAF(mutation.csv,
                        gene.symbol.col = "Hugo_Symbol",
                        variant.class.col = "Variant_Classification",
                        protein.change.col = "amino_acid_change",
                        sep = ",")  # column-separator of csv file


plot.options <- g3Lollipop.options(
  # Chart settings
  chart.width = 600,
  chart.type = "pie",
  chart.margin = list(left = 30, right = 20, top = 20, bottom = 30),
  chart.background = "#d3d3d3",
  transition.time = 300,
  # Lollipop track settings
  lollipop.track.height = 200,
  lollipop.track.background = "#d3d3d3",
  lollipop.pop.min.size = 1,
  lollipop.pop.max.size = 8,
  lollipop.pop.info.limit = 5.5,
  lollipop.pop.info.dy = "0.24em",
  lollipop.pop.info.color = "white",
  lollipop.line.color = "#a9A9A9",
  lollipop.line.width = 3,
  lollipop.circle.color = "#ffdead",
  lollipop.circle.width = 0.4,
  lollipop.label.ratio = 2,
  lollipop.label.min.font.size = 12,
  lollipop.color.scheme = "dark2",
  highlight.text.angle = 60,
  # Domain annotation track settings
  anno.height = 16,
  anno.margin = list(top = 0, bottom = 0),
  anno.background = "#d3d3d3",
  anno.bar.fill = "#a9a9a9",
  anno.bar.margin = list(top = 4, bottom = 4),
  domain.color.scheme = "pie5",
  domain.margin = list(top = 2, bottom = 2),
  domain.text.color = "white",
  domain.text.font = "italic 8px Serif",
  # Y-axis label
  y.axis.label = "# of PIK3CA gene mutations",
  axis.label.color = "#303030",
  axis.label.alignment = "end",
  axis.label.font = "italic 12px Serif",
  axis.label.dy = "-1.5em",
  y.axis.line.color = "#303030",
  y.axis.line.width = 0.5,
  y.axis.line.style = "line",
  y.max.range.ratio = 1.1,
  # Chart title settings
  title.color = "#303030",
  title.text = "PIK3CA gene (customized chart options)",
  title.font = "bold 12px monospace",
  title.alignment = "start",
  # Chart legend settings
  legend = TRUE,
  legend.margin = list(left=20, right = 0, top = 10, bottom = 5),
  legend.interactive = TRUE,
  legend.title = "Variant classification",
  # Brush selection tool
  brush = TRUE,
  brush.selection.background = "#F8F8FF",
  brush.selection.opacity = 0.3,
  brush.border.color = "#a9a9a9",
  brush.border.width = 1,
  brush.handler.color = "#303030",
  # tooltip and zoom
  tooltip = TRUE,
  zoom = TRUE
)


```

```{r amino_acid_mutation_plot}

g3Lollipop(mutation.dat,
           gene.symbol = mutation_gene_hugo,
           protein.change.col = "amino_acid_change",
           btn.style = "blue", # blue-style chart download buttons
           plot.options = plot.options,
           output.filename = "customized_plot")
```



<p> </p>

<div class="pd22">
GenVisR is  focused on supporting of genomic visualizations. The library waterfall of GenViSR answers the question  what mutations are present in the sample. This type plot displays  in a heatmap-like structure with rows and columns denotating genes and samples. </div>


The goal is to investigate Bioconductor package GenVisR and to apply it on a clinical dataset. The plan is to  conduct deeper research on  GenVisR, develop new data analysis techniques, and make contributions to the fields of mutational study in  biology and medicine.
</p>

<p> Dataset supposed to come from Ensemble  database



