# Data managment principles




<div class="azure"> 


Applying the Guerrilla analytics framework to your own project.

    Look at your RStudio project that you created for the DAUR-II assignments
    download your folders (not the HUGE datafiles and kraken-database!) and rearrange your project according the Guerilla principles explained above
    for the data… folders: provide only a readme.txt with alle the info you would normally put there + where the data is stored
    Provide study documentation and meta data.
    Use the {fs} package to share a screenshot of your folder tree in your portfolio, look at: for more info on how to use the {fs} package.

hint: you had 3 datasets within the RNAseq project, one of them being the exam project, and 2 datasets within the metagenomics project.

    Now clean up your work environment for this course (workflows) and the parallel course in DSFB2 (projecticum). Set up a folder structure that will accomodate future plans and collaboration on the projecticum. Provide readme-files or comments within the code where needed.

For the projecticum folder, make sure you do this together with your lab partner.

</div>





<h5> The portfolio directory: </h5>

```{r, warning=FALSE, message=FALSE}

library(fs)

```



```{r}

                              
fs::dir_tree("C:/Users/Mijn-PC/Documents/hu_courses/worfklow_ds_2")
```



<h5> The Daur-2 directory: </h5>

```{r}


fs::dir_tree("C:/Users/Mijn-PC/Documents/hu_courses/Daur2")
```



<h5> The projecticum  directory: </h5>

```{r}


fs::dir_tree("C:/Users/Mijn-PC/Documents/hu_courses/projecticum")
```


<h5> The portfolio  directory: </h5>

```{r}


fs::dir_tree("C:/Users/Mijn-PC/Documents/hu_courses/portfolio")
```


Include readme to the projecticum dataset


<object data="assets/vignettes/readme_dataset.pdf" type="application/pdf" width="100%" height="750px">
  <embed src="assets/vignettes/readme_dataset.pdf">
  </object>



