#  Working with relational databases


<h3>  Portfolio assignment 7.Relational databases  </h3>


```{r, warning=F, message=F}
library(tidyverse)
library(ggplot2)
library("dslabs")
```


<div class="azure" >Assignment

    Load the flu (in the github repo for this course: https://github.com/DataScienceILC/tlsc-dsfb26v-20_workflows go to ./data/flu_data.csv), the dengue (“./data/dengue_data.csv) and the gapminder ({dslabs} package) into three separate dataframes in R

    Check if they are in the right shape. Is the data in the ‘tidy’ format? If not change the format to ‘tidy’

    Change the country and date variables of the three tables so that they coincide in terms of data type, class and values

    Store the three tables as separate (so six in total) .csv and .rds files.

    In Dbeaver create a new PostgreSQL database “workflowsdb”

    Using RPostgreSQL, insert the tables into the database.

    Inspect the contents of the tables with SQL (in DBeaver) and save the SQL script.

    Inspect the contents of the tables with dplyr (in R) and save a RMarkdown showing what you are doing.

    Load the gapminder data in R and change the dataframe in such as way that you could join it to dengue and flue.

    Save this clean gapminder data in the “workflowsdb” database

    Perform some joins (your choice) with SQL (can be done in DBeaver or with dplyr.

    Generate a joined table, and export this from the database to R.

    Show some descriptive statistics with this table, and at least 3 visualisations using ggplot2.

    Write a short report to show at least te actions listed in this assignment in a Rmd file for your portfolio. Include pictures and provide text explaining and showcasing your skills.
 </div>


<h3> read in Flu  dataset </h3>


```{r}
flu <-   read.csv("data_raw/les5/flu_data.csv", skip=11)
colnames(flu)

```

<h3> read in  Dengue dataset  </h3>

```{r}
dengue <-  read.csv("data_raw/les5/dengue_data.csv", skip=11)

colnames(dengue)
```

<h3>  Read in dslabs gapminder dataset  </h3>


```{r}
library("dslabs")

head(gapminder)
str(gapminder)
```

<h3> Processing of  FLU dataset   </h3>


```{r}
flu_longer <-  pivot_longer(flu, cols= -Date,  names_to = "country", values_to ="flu_cases", values_drop_na = TRUE)
head(flu_longer)
```
 <h3>  Fromat date, with year extraction </h3>

```{r}
flu_longer$Date <- parse_integer( format(as.Date(flu_longer$Date), format="%Y") )
head(flu_longer$Date)

head(flu_longer)
```

<h3> Save unique rows </h3>


```{r}
flu_longer2 <-  flu_longer %>%   group_by(Date, country)  %>%
  mutate( sum_cases = sum(flu_cases))

#View(flu_longer2)

head( flu_longer2)

colnames( flu_longer2)[1] <-  "year"

colnames(flu_longer2)


flu_longer_db <-  flu_longer2 %>%  select (year, country, sum_cases) %>%  unique()

```






<h3>  Processing of  Dengue dataset  </h3>

```{r}
dengue_longer <-  pivot_longer(dengue, cols = -Date, names_to = "country", values_to = "dengue_cases",
                               values_drop_na = TRUE)

head(dengue_longer)

dengue_longer$Date <-  parse_integer( format(as.Date(dengue_longer$Date), format="%Y") )


head(dengue_longer)

dengue_longer <- dengue_longer %>%  group_by(country, Date) %>%  mutate(dengue_sum_cases = sum(dengue_cases))

head(dengue_longer)

colnames(dengue_longer)[1]  <-  "year"

colnames(dengue_longer)

dengue_db <-  dengue_longer %>%  select (year, country, dengue_sum_cases) %>%  unique()

dengue_db


```


<h3> export  tables to csv and rds files  </h3>



```{r, eval=F}
write.csv(flu_longer_db, "output/flu.csv")
write.csv( dengue_db, "output/ dengue.csv")
write.csv( gapminder, "output/gapminder.csv")

saveRDS(flu_longer_db, "output/flu.rds")
saveRDS( dengue_db, "output/dengue.rds")
saveRDS(gapminder, "output/gapminde.rds")


```


<h3> Make connection to Postgres DB  </h3>


```{r}
library('RPostgreSQL')
library("remotes")
library("RPostgres")

con <- dbConnect(RPostgres::Postgres(),
                 dbname = "workflowsdb",
                 host="localhost",
                 port="5432",
                 user="postgres",
                 password="admin")
```

<h3> write tables to DB  </h3>


```{r, eval=F}
dbWriteTable(con, "flu",flu_longer_db)

dbWriteTable(con, "dengue", dengue_db)

dbWriteTable(con, "gapminder", gapminder)

dbListTables(con)
```

inspect the contents of the tables with SQL


```{sql connection=con}
select * from flu limit 10;
```

```{sql connection=con}
select * from dengue limit 10;
```


<h3> Perform sql joins in DB: </h3>

1. inner join flu and dengue:

```{sql connection=con, eval= FALSE}

CREATE TABLE flu_joined_dengue
AS 
SELECT  dengue.country, dengue.year , flu.sum_cases,  dengue.dengue_sum_cases   from  dengue  join flu  on   dengue.year = flu.year and dengue.country = flu.country;
```


```{sql connection=con}
select * from flu_joined_dengue limit 10;
```

<h3> Left join on gapminder to combined flu-dengue  </h3>


```{sql connection=con}
select gapminder.country, gapminder.year, gapminder.infant_mortality , gapminder.life_expectancy , gapminder.continent, flu_joined_dengue.sum_cases ,flu_joined_dengue.dengue_sum_cases  from gapminder   left join  flu_joined_dengue  on flu_joined_dengue.year= gapminder.year and flu_joined_dengue.country =gapminder.country ;

```

<h3> Inspect data with R </h3>

```{r}


flu_db <- tbl(con, "flu")
head(flu_db)

```

```{r}
debgue_db <- tbl(con, "dengue")
head(debgue_db)
```

```{r}
gapminder_db <- tbl(con, "gapminder")
head(gapminder_db)
```



<h3> Descriptive statistics </h3>


1. Join: flu and dengue"


```{r}
join2 <-  read.csv("data/les5/join2.csv")

colnames(join2)[3] <-  "cases_flu"
colnames(join2)[4] <-  "cases_dengue"

head(join2)


join2 %>%  filter(country %in% c("Bolivia", "Brazil", "Argentina")  )  %>%
  ggplot(aes(x=year, y=cases_dengue, color=country)) +
  geom_line(size=2.5)+
  labs(title = "Comparing of dengue cases in Argentina, Bolivia and Brazil" ) 


join2 %>%  filter(country %in% c("Bolivia", "Brazil", "Argentina")  )  %>%
  ggplot(aes(x=year, y=cases_flu, color=country)) +
  geom_line(size=2.5)+
    labs(title = "Comparing of flu cases  in Argentina, Bolivia and Brazil" ) 


```



```{r}


summary_table <- join2 %>%  group_by(country) %>%  summarise( max_flu=max(cases_flu), max_dengue=max(cases_dengue)  )



join2_longer <-join2 %>%  pivot_longer( cols = c("cases_flu","cases_dengue"),
                   names_to = "group", values_to ="cases" )
head(join2_longer)



# bar chart flu vs dengue


summary_table <- join2 %>%  group_by(country) %>%  summarise( max_flu=max(cases_flu), max_dengue=max(cases_dengue)  )

summary_table
```


```{r}




summary_table_polt  <-  summary_table %>% pivot_longer(cols=c('max_flu', 'max_dengue'),
                           names_to='max_flu_dengue',
                           values_to='frequentie')

head(summary_table_polt)

ggplot (summary_table_polt, aes(x= max_flu_dengue, y=log(frequentie), fill=country )) +
  geom_bar ( stat="identity", position = position_dodge(width = 0.5))+
  labs(title="Maximum cases Flu and Dengue cases per country")+
  ylab("Frequentie, log based " )+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = frequentie), size = 3, vjust = -0.4,position = position_dodge(width = 1) )



ggplot (summary_table_polt, aes(x= country, y=log(frequentie), fill=max_flu_dengue )) +
  geom_bar ( stat="identity", position = position_dodge(width = 0.5))+
  labs(title="Maximum cases  Flu and  Dengue per country")+
  ylab("Frequentie, log based " )+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = frequentie), size = 3, vjust = -0.4,position = position_dodge(width = 1) )
```


```{r, echo=FALSE}
maxDengueArgenitina <-  summary_table_polt$frequentie[summary_table_polt$country=="Argentina" & summary_table_polt$max_flu_dengue=="max_dengue"]

maxDengueBolivia <-  summary_table_polt$frequentie[summary_table_polt$country=="Bolivia" & summary_table_polt$max_flu_dengue=="max_dengue"]


maxDengueBrazil <-  summary_table_polt$frequentie[summary_table_polt$country=="Brazil" & summary_table_polt$max_flu_dengue=="max_dengue"]
 

maxDengueMexico <-  summary_table_polt$frequentie[summary_table_polt$country=="Mexico" & summary_table_polt$max_flu_dengue=="max_dengue"]

```




<p> Maximum dengue cases in Argenitina is : `r maxDengueArgenitina` </p>
<p> Maximum flu cases in Mexico is : `r maxDengueMexico`  </p>
<p> Maximum flu cases in Brazil is:  `r maxDengueBrazil`  </p>
<p> Maximum flu cases in Bolivia is :`r maxDengueBolivia`  </p>

<p> There are thousands of times more cases of flu than dengue in all existing areas </p>



<h3> Left join : gapminder and combined flu and dengue </h3>


```{r}
join1 <-  read.csv("data/les5/left_join_gapminder.csv")
summary(join1)


join1 %>% 
  ggplot(aes(x=log(sum_cases) , y=dengue_sum_cases, color=continent)) +
  geom_point()+
    labs(title = "Scatterplot of flu and dengue cases ", y= "flu cases, natural log", x="dengue cases"  )
  

```



<p> There is  no positive or negative  correlation between flu cases and  dengue cases in all given countries (continent  America ) </p>
<p> 
Flu cases are several thousand times higher than dengue cases , with a maximum of 70,000 flue cases in Mexica. While the maximum number of dengue cases is 16, in Argentina  </p>
